---
title: "ggvis cookbook in plotly kitchen"
output:
  html_document:
    toc: true
---

<!--
%\VignetteEngine{knitr::rmarkdown}
%\VignetteIndexEntry{ggvis cookbook}
-->

```{r, echo = FALSE, message = FALSE}
library(knitr)
opts_chunk$set(comment = "#>", error = FALSE, tidy = FALSE)
opts_chunk$set(fig.width = 2, fig.height = 1.25, dpi = 100)
```

The following examples show you how to create a selection of common graphics with ggvis. First, load ggvis and dplyr:

```{r, message = FALSE}
library(plotly)
library(ggvis)
library(dplyr)
```

## Scatterplots

We'll use the built-in `mtcars` data set, and look at two columns of interest, `mpg`, and `wt`:

```{r}
# The first few rows of mtcars
head(mtcars)
#mtcars %>% ggvis(~wt, ~mpg) %>% layer_points()
mtcars %>% plot_ly(x=~wt, y=~mpg) %>% add_markers()
```

Smaller points, a different shape, a different outline (stroke) color, and empty fill:

```{r}
mtcars %>% 
  #ggvis(~wt, ~mpg) %>% 
  plot_ly(x=~wt, y=~mpg) %>%
  #layer_points(size := 25, shape := "diamond", stroke := "red", fill := NA)
  add_markers(marker=list(size=25,symbol="diamond-open",color='red'))
```


## Regression lines

Adding a smooth line

```{r, message = FALSE}
mtcars %>% 
  #ggvis(~wt, ~mpg) %>%
  plot_ly(x=~wt, y=~mpg) %>%
  #layer_points() %>%
  add_markers() %>%
  #layer_smooths()
  add_lines(y=~fitted(loess(mpg~wt))[order(wt)],x=~sort(wt))
```

With a linear model, and 95% confidence interval for the model:

**todo**

```{r, message = FALSE}
mtcars %>% 
  ggvis(~wt, ~mpg) %>%
  layer_points() %>%
  layer_model_predictions(model = "lm", se = TRUE)
```


## Scatter plots with grouping

Coloring points by a variable:

```{r, message = FALSE}
mtcars %>% 
  #ggvis(~wt, ~mpg) %>% 
  plot_ly(x=~wt, y =~mpg) %>%
  #layer_points(fill = ~factor(cyl))
  add_markers(color=~factor(cyl))
```

Coloring points, and adding a smoother for each group. The grouping variable (which is applied before the `transform_smooth` is calculated) must be specified with `group_by()`:

**todo**

```{r, message = FALSE}
mtcars %>% 
  #ggvis(~wt, ~mpg, fill = ~factor(cyl)) %>% 
  plot_ly(x=~wt, y=~mpg, color=~factor(cyl)) %>%
  #layer_points() %>% 
  add_markers()# %>%
  #group_by(cyl) %>% 
  #layer_model_predictions(model = "lm")
```

## Bar graphs

We'll use the built-in `pressure` data set for these examples:

```{r}
# The first few rows
head(pressure)
```

When the variable on the x axis is continuous (e.g., numeric or date-time):

```{r, message = FALSE, fig.width = 4}
pressure %>% 
  #ggvis(~temperature, ~pressure) %>%
  plot_ly(x=~temperature, y=~pressure, type="bar")
  #layer_bars()
  
```

It's possible to specify the width of the bars:

```{r, message = FALSE, fig.width = 4}
pressure %>% 
  #ggvis(~temperature, ~pressure) %>%
  plot_ly() %>%
  add_trace(x=~temperature, y=~pressure, type="bar") %>%
  #layer_bars(width = 10)
  layout(bargap=0.6)
```

When the variable on the x axis is categorical (e.g., factor or character):

```{r, message = FALSE, fig.width = 4}
# First, modify the pressure data set so that the x variable is a factor
pressure2 <- pressure %>% mutate(temperature = factor(temperature))

pressure2 %>%
  #ggvis(~temperature, ~pressure) %>%
  plot_ly(x=~temperature, y=~pressure, type="bar")
  #layer_bars()
```


## Line graphs

```{r, message = FALSE}
pressure %>%
  #ggvis(~temperature, ~pressure) %>%
  plot_ly(x=~temperature, y=~pressure) %>%
  add_lines()
  #layer_lines()
```

Lines with points:

```{r, message = FALSE}
pressure %>%
  #ggvis(~temperature, ~pressure) %>%
  plot_ly(x=~temperature, y=~pressure)
  #layer_points() %>% 
  #layer_lines()
```

## Histograms

We'll use the built-in `faithful` data set for these examples:

```{r}
# The first few rows
head(faithful)
```

Basic histogram:

```{r, message = FALSE}
faithful %>%
  #ggvis(~eruptions) %>%
  plot_ly(x=~eruptions, type="histogram")
  #layer_histograms()
```


The bin selection can be controled by specifying `width` and at most one of `center` or 
`boundary` of one of the bins.  `boundary` and `center` may be outside the range of 
the data.

**todo**
**more informative exploration of histogram**

```{r, message = FALSE}
faithful %>%
  #ggvis(~eruptions) %>%
  plot_ly( x=~eruptions ) %>%
  #layer_histograms(width=0.5, boundary=0)
  add_trace(
    type="histogram",
    xbins=list(
      start=0,
      end=10,
      size=0.5
    ),
    autobinx=FALSE
  )

# todo
#  easy way to get ticks at center of bins
#  and match our xbins settings
faithful %>%
  #ggvis(~eruptions) %>%
  plot_ly( x=~eruptions ) %>%
  #layer_histograms(width=0.5, center=0)
  add_trace(
    type="histogram",
    xbins=list(
      start=0,
      end=10,
      size=0.5
    ),
    autobinx=FALSE
  ) %>%
  layout(
    xaxis=list(
      tickmode="array",
      tickvals=seq(0,10,0.5) + 0.25
    )
  )
  
```

Modify the fill color and bin width, and add titles for the axes, since the automatic titles aren't very informative:

**todo**
**add stroke color black on bars**

```{r, message = FALSE}
faithful %>%
  #ggvis(~eruptions, fill := "#fff8dc") %>%
  plot_ly(x=~eruptions, marker=list(color="#fff8dc")) %>%
  add_trace(
    type="histogram",
    autobinx=FALSE,
    xbins=list(
      start=0,
      end=10,
      size=0.25
    )
  )
  #layer_histograms(width = 0.25)
```

By default, when the number of integer values is small, bins will 
be centered at integers and have a width of 1:

**todo**
**add description of defaults**

```{r, message = FALSE}
data(cocaine, package="ggvis")
cocaine %>% 
  #ggvis(~month, fill := "#fff8dc") %>%
  plot_ly(x=~month, marker=list(color="#fff8dc")) %>%
  #layer_histograms() %>%
  add_trace(type="histogram") %>%
  #add_axis("x", title = "month") %>%
  #add_axis("y", title = "count")
  layout(
    xaxis=list(title="month"),
    yaxis=list(title="count")
  )
```

This can be forced with 
```{r, message = FALSE}
cocaine %>%
  #ggvis(~month, fill := "#fff8dc") %>%
  plot_ly(x=~month, marker=list(color="#fff8dc")) %>%
  #layer_histograms(width = 1, center = 0) %>%
  add_trace(
    type="histogram",
    autobinx=FALSE,
    xbins=list(
      start=0,
      end=12,
      size=1
    )
  ) %>%
  #add_axis("x", title = "month") %>%
  #add_axis("y", title = "count")
  layout(
    xaxis=list(
      title="month",
      tickmode="array",
      tickvals=0:12+0.5,
      ticktext=0:12,
      ticks="outside"
    ),
    yaxis=list(title="count"),
    bargap=0.25
  )
```

## Box plots

```{r message=FALSE, warning=FALSE}
mtcars %>%
  #ggvis(~factor(cyl), ~mpg) %>%
  plot_ly(x=~factor(cyl),y=~mpg) %>%
  add_trace(type="box")
  #layer_boxplots()
```

